---
title: "R Notebook"
output: html_notebook
---

ts analysis app drafting
goal - input a univariate time series and return visualizations, and forecast metrics for different TS algorithms


```{r}
df <- read.csv(file.choose())
names(df) <- c('units', 'date')
df$date <- df$date %>% as.Date()
```

```{r}
#algorithms
library(rmdformats)
library(tidyverse)  # data manipulation
library(data.table)
#library(tswge)  # Time series package
library(tseries)  # for Dickey-Fuller test 
library(orcutt)  # for Cochrane-Orcutt test
library(formattable)  # for table formatting
library(GGally)
library(astsa)
library(nnfor)
library(dplyr)
library(ggplot2)
library(changepoint)
library(date)
library(R.devices)

### A series of functions used in a time series analysis dashboard capstone project for MSDS at SMU ###

#Equal Means
#accept df object and return forecasts n_ahead
em_ts <- function(n_ahead, target_df){
  return(rep((mean(target_df$units)),  n_ahead))}

#bootleg AIC5 with no tswge
#target_vector <- target_df$STD_Cases 
aic_boot <- function(target_vector){
  pqc <- expand.grid(0:5, 0:2)
  aic_holder <- vector()
  p_holder <- vector()
  q_holder <- vector()
  for(i in 1:length(pqc$Var1)) {
    temp_model <- arima(target_vector, order = c(pqc[i, 1], 0, pqc[i, 2]))
    #aic_holder <- append(aic_holder, temp_model$aic) #base AIC with k = 2?
    aic_holder[i] <- AIC(temp_model, k = log(length(target_vector)))
    p_holder <- append(p_holder, pqc[i, 1])
    q_holder <- append(q_holder, pqc[i, 2])}
  aic_results <- data.frame(p = p_holder, q = q_holder, aic = aic_holder)
  aic_results <- aic_results[order(aic_results$aic), ]
  return(head(aic_results, 1))
}

aic_boot2 <- function(target_vector){
  pqc <- expand.grid(0:5, 0:2)
  aic_holder <- vector()
  p_holder <- vector()
  q_holder <- vector()
  for(i in 1:length(pqc$Var1)) {
    temp_model <- arima(target_vector, order = c(pqc[i, 1], 0, pqc[i, 2]), method = 'ML')
    #aic_holder <- append(aic_holder, temp_model$aic) #base AIC with k = 2?
    aic_holder[i] <- AIC(temp_model, k = log(length(target_vector)))
    p_holder <- append(p_holder, pqc[i, 1])
    q_holder <- append(q_holder, pqc[i, 2])}
  aic_results <- data.frame(p = p_holder, q = q_holder, aic = aic_holder)
  aic_results <- aic_results[order(aic_results$aic), ]
  return(head(aic_results, 1))
}

#AR model bootleg
ar_ts_boot <- function(n_ahead, df){
  aic_temp <- aic_boot2(df[, 1])
  if (aic_temp$p == 0){
    newphi = 1}
  else{newphi = aic_temp$p}
  model_temp <- arima0(df$units, order = c(newphi, 0, 0), method = 'ML') #note q is zero for AR model
  pred_temp <- invisible(predict(model_temp, n_ahead))
  return(pred_temp$pred)
}

arma_ts_boot <- function(n_ahead, df){
  aic_temp <- aic_boot2(df[, 1])
  if (aic_temp$p == 0){
    newphi = 1}
  else newphi = aic_temp$p #error if inside curly brackets??? 
  model_temp <- arima0(df$units, order = c(aic_temp$p, 0, newphi)) #note q is not zero for ARMA model
  pred_temp <- predict(model_temp, n_ahead) %>% invisible()
  return(pred_temp$pred)
}

#ARI no tswge
ari_ts_boot <- function(n_ahead, df){
  diff1 <- diff(df$units)
  aic_temp <- aic_boot2(diff1)
  if (aic_temp$p == 0){
    newphi = 1}
  else{newphi = aic_temp$p}
  pred_temp <- predict(arima0(df$units, order = c(newphi, 1, 0)), n.ahead = n_ahead)
  return(pred_temp$pred)
}

#no tswge
arima_ts_boot <- function(n_ahead, df){
  diff1 = diff(df$units)
  aic_temp <- aic_boot2(diff1)
  if (aic_temp$p == 0){
    newphi = 1}
  else{newphi = aic_temp$p}
  pred_temp <- predict(arima0(df$units, order = c(newphi, 1, aic_temp$q)), n.ahead = n_ahead)
  return(pred_temp$pred) #consider replacing df$units with df[, 1] to be name agnostic, if there's a way to guarantee the df from csv is column 1 data column 2 date
}

#bootleg ARI_S12
ari_s12_ts_boot <- function(n_ahead, df){
  diff1 <- diff(df$units)
  aic_temp <- aic_boot2(diff1)
  if (aic_temp$p == 0){
    newphi = 1
  } else {
    newphi = aic_temp$p
  } 
  pred_temp <- predict(arima0(df$units, order = c(newphi, 0, 0), seasonal = c(0, 0, 12)), n.ahead = n_ahead)
  return(pred_temp$pred) 
  }

#bootleg ARIMA_S12
arima_s12_ts_boot <- function(n_ahead, df){
  diff1 <- diff(df$units)
  aic_temp <- aic_boot2(diff1)
  if (aic_temp$p == 0){
    newphi = 1
  } else {
    newphi = aic_temp$p
  } 
  pred_temp <- predict(arima0(df$units, order = c(newphi, 0, aic_temp$q), seasonal = c(0, 0, 12)), n.ahead = n_ahead)
  return(pred_temp$pred) }
#arima_s12_ts_boot(5, target_df) #+2% off from the tswge function

#RF function
library(ranger)
library(randomForest)
library(caret)
#with no graph
rf_ts <- function(n_ahead, df){ #can add in df with additional variables, ex units, date, sales
  #target_df$date <- as.Date(target_df$date, '%m/%d/%Y')
  tlen = length(df[, 1])
  sample_train <- df[1:(tlen-n_ahead), ]
  sample_test <- df[(tlen-n_ahead +1):tlen, ]
  #sample_test[, 3][is.na(sample_test[, 3])] <- 0 #check 3rd var for NA, convert to zero if found
  rf1 <- randomForest(units ~ date , data = sample_train, na.action = na.exclude)
  #not sure how to make this adaptive, would like to have the variables be units or col1 vs relationship to all other columns, come back to this later, leave as univariate for now
  pred1 = predict(rf1, newdata = sample_test)
  rf_ase <- mean((sample_test$units - pred1)^2)
  predictions <- data.frame(date = sample_test$date, forecast = pred1)
  results = c(predictions, data.frame(ASE = rf_ase))
  #return(results)
  return(results$forecast)
}#note, RandomForest doesn't work so well with only 1 predictor, esp if linear progression like in example data



#nn complete
nnc <- function(target_df, n_ahead, reps, hd_vector, display){
  #date_line <- as.Date(target_df$date, '%m/%d/%Y') #add date back in after it works
  tlen <- length(target_df[, 1])
  sample_train <- target_df[1:(tlen-n_ahead), ]
  sample_test <- target_df[(tlen-n_ahead +1):tlen, ]
  
  ts_train<- as.ts(sample_train[, 1])
  fit_mlp <- mlp(ts_train, reps = reps, hd = hd_vector)
  
  mlp_pred <- forecast(fit_mlp, h = n_ahead)
  mlp_ase <- mean((sample_test[, 1] - mlp_pred$mean)^2)
  
  #r1 <- ggplot() + 
  #  geom_line(aes(x = sort(target_df$date), y = target_df$STD_Cases), col = 'black') +
  #  geom_line(aes(x = sort(target_df$date), y = c(rep(NA, (tlen-n_ahead)), mlp_pred$mean)), col = 'red') +
  #  ggtitle(paste('MLP', n_ahead, 'Month ASE: ', mlp_ase), subtitle =paste(target_df$Product[1], 'Purchases By Cust ID: ', target_df$Customer_ID[1]))
  
  #if(display == TRUE){
  #  print(fit_mlp)
  #  plot(fit_mlp)
  #  show(r1)}
  
  forecasts <- data.frame(forecast = mlp_pred$mean)
  return(c(forecasts, mlp_ase))
}

#function that accepts function for collated rolling ase 1:12 based on trainingSize ect, so i can just swap ar_ts with rf_ts and reuse the same evaluation matrix 
ase_gen <- function(target_df, FUNCTION){
  #FUNCTION = ar_ts
  j = 12 #n ahead
  ntotal <- length(df[, 1])#+1
  trainingSize = round(ntotal*.7)
  #60 #will cause problems with the shorter datasets if left at flat 60
  temp_roll_ase = c() #holder
  temp_ase_matrix = matrix(0, ncol = 12, nrow = length(1:(ntotal-(trainingSize + j) + 1))) #preallocation
  
  for(k in 1:(ntotal-(trainingSize + j) + 1)){
    # print(substitute(FUNCTION)) #use subsitute to return function name as string
    if(substitute(FUNCTION) == 'nnc'){ #use subsitute to return function name as symbol/match
      temp_forecasts <- FUNCTION(target_df[k:(k+(trainingSize-1)),], j, 5, c(10, 10, 10), FALSE)
      for(i in 1:12){
        wstart <- trainingSize + k 
        wend <- trainingSize + k + i -1
        #print(paste(wstart, ':', wend))
        temp_ase_matrix[k, i] <- mean((target_df[, 1][(wstart:wend)]-temp_forecasts$forecast[1:i])^2)
      }}
    else{
      temp_forecasts <- FUNCTION(j, target_df[k:(k+(trainingSize-1)), ])#feed ar_ts dataframe with shifting rows
      for(i in 1:12){
        wstart <- trainingSize + k 
        wend <- trainingSize + k + i -1
        #print(paste(wstart, ':', wend))
        temp_ase_matrix[k, i] <- mean((target_df[, 1][(wstart:wend)]-temp_forecasts[1:i])^2)
      }}
  }
  #print(k)
  
  closeAllConnections()
  for(i in 1:12){
    temp_roll_ase[i] <-mean(temp_ase_matrix[, i])}
  return(temp_roll_ase)
  #return(temp_ase_matrix)
}
#have to sink/closeallconnections the function to hide tswge table output
```

initial vis and diagnostics
```{r}
library(ggplot2)
v0 <- ggplot(data = df) +
  geom_line(aes(x = date, y = units)) +
  ggtitle('univariate time series')
v0

pacf(df)
```

#individual plots for each alg??
```{r}
n_ahead <- 6 #slider input here
train_df <- df[1:(length(df[, 1])-(n_ahead)), ]
test_df <- df[(length(df[, 1])-n_ahead+1):length(df[, 1]), ]
pred <- em_ts(n_ahead, train_df)
pred_df <- data.frame('units' = pred, 'date' = test_df$date)
v1 <- ggplot() +
  geom_line(data = train_df, aes(x = date, y = units, col = 'a'))+
  geom_line(data = pred_df, aes(x = date, y = units, col = 'b')) + 
  geom_line(data = test_df, aes(x = date, y = units, col = 'c')) +
  scale_color_discrete(name = ' ', labels = c('original', 'prediction', 'actual'))
v1

```
a bit more adaptive, specify function, df, n_ahead -> graph
eventual flow needs to be char -> function ->
```{r}
fitcheck <- function(n_ahead, df, FUNCTION_STRING){
  train_df <- df[1:(length(df[, 1])-(n_ahead)), ]
test_df <- df[(length(df[, 1])-n_ahead+1):length(df[, 1]), ]
alg <- match.fun(FUNCTION_STRING)
#exception for nnc
if(FUNCTION_STRING == 'nnc'){
  x <- alg(df, n_ahead, 20, c(5, 5, 5), FALSE)
  pred <- x[[1]][1:6] #needs to be flat list for consistency
}
else{
  pred <- alg(n_ahead, train_df)
}

pred_df <- data.frame('units' = pred, 'date' = test_df$date)
vx <- ggplot() +
  geom_line(data = train_df, aes(x = date, y = units, col = 'a'))+
  geom_line(data = pred_df, aes(x = date, y = units, col = 'b')) + 
  geom_line(data = test_df, aes(x = date, y = units, col = 'c')) +
  scale_color_discrete(name = ' ', labels = c('original', 'prediction', 'actual')) +
  ggtitle(paste0('Prediction Algorithm: ', FUNCTION_STRING), subtitle = paste0('Check fit on: ', n_ahead, ' [time intervals]'))
return(vx)
}
```

```{r}
#basic execution
#have a slider for n_ahead
#have a menu/buttons for FUNCTION? list of functions as strings, convert to function with match.fun
#nevermind, match.fun simply assigns the listed function to a new identifier
#input df on earlier screen?? probably need a check for that
#also add ASE? or another display for that? 

n_ahead <- 12
#fitcheck accepts a string matching a function name
v1 <- fitcheck(n_ahead, df, 'nnc')
v1
```


generate all forecast data and store in df and plot????
```{r}
any_backcast <- function(n_ahead, df, FUNCTION_STRING){
  train_df <- df[1:(length(df[, 1])-(n_ahead)), ]
test_df <- df[(length(df[, 1])-n_ahead+1):length(df[, 1]), ]
alg <- match.fun(FUNCTION_STRING)
#exception for nnc
if(FUNCTION_STRING == 'nnc'){
  x <- alg(df, n_ahead, 20, c(5, 5, 5), FALSE)
  pred <- x[[1]][1:6] #needs to be flat list for consistency
}
else{
  pred <- alg(n_ahead, train_df)
}

pred_df <- data.frame('units' = pred, 'date' = test_df$date)
return(pred_df)
}
#use this to iteratively generate predictions keyed to specified dates
```

```{r}
n_ahead <- 12 #slider input hooks here
alg_list <- c('em_ts', 'ar_ts_boot', 'ari_ts_boot', 'ari_s12_ts_boot', 'arima_ts_boot', 'arima_s12_ts_boot', 'nnc', 'rf_ts')

f_frame <- data.frame('n_ahead' = seq(1, n_ahead, 1)) #prep frame
for(a in alg_list){
  temp_pred <-  any_backcast(n_ahead, df, a)
  names(temp_pred) <- c(paste(a), 'date') #leave date column in for diagnostics
  f_frame <- cbind(f_frame, temp_pred)
}

f_frame <- f_frame[, !duplicated(colnames(f_frame))] #remove duplicated date columns leave off for sanity check on date matchups
f_frame <- f_frame[, -1] #remove n_ahead placeholder
f_frame <- subset(f_frame, select = c(2, 1, 3:8)) #move date to the first column
f_frame #results frame

f_melt <- melt(f_frame, id.vars = 'date')
#graph f_frame
v_all <- ggplot()+
  #geom_line(data = df, aes(x = date, y = units)) +
  geom_line(data = df[(length(df$units)-n_ahead):length(df$units), ], aes(x = date, y = units))+
  geom_line(data = f_melt, aes(x = date, y = value, col = variable)) +
  ggtitle(paste('Most Recent [', n_ahead, '] Time Units vs [', length(alg_list), '] TS Algorithms'))
v_all
```
room to put ase on a seperate data table, maybe with highlighted row that has the lowest ASE

```{r}
#generate ASE per method (simple, single run per method)
#Average Squared Error (higher is less accurate*)
n_ahead <- 12
actual_data <- df[(length(df$units)-n_ahead+1):length(df$units), ] #the data to compare to
#use f_frame from before? or might have to calc again depending on page?


ASE <- data.frame()
for(c in seq(2, ncol(f_frame)-1)){ #requires date to be the first column to skip calc
  temp_ase <- (actual_data$units - f_frame[, c])^2 %>% mean()
  ASE <- rbind(ASE, temp_ase)
}
row.names(ASE) <- names(f_frame)[seq(2, ncol(f_frame)-1)]
names(ASE) <- 'ASE'
print(ASE)

#add other accuracy metrics here later? 

```

 
goals - 
- generate forecast data with slider/input from ui for months/time units ahead
- selective display of models used, checkbox/radar buttons for showing original data vs arma, arima, rf, ect
  - would likely require collation of all models run to n length in df, then selecting specific slices of the frame to display
- interactive tool tips for mouseover of ts to show point values, date and units for actual/prediction(s)
- some sort of indicator that calculations are ongoing
panels
1 - diagnostics
2 - looking backwards/fitting algorithm - show ase calcs
3 - looking forwards -> apply alg WITH/AFTER original data






