---
title: "WF scraper test1"
author: "Jonathan Tan"
date: "8/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
goal: create an adaptive program that takes the name of a wf weapon
and scrapes the wf wiki page for stats, possible pictures? 

tools: rvest, html targeting

brainstorming:
retrieval mechanism should navigate to a webpage, 
warframe.fandom.com/wiki/[weapon name here]
and retrieve 
mastery rank
slot
type
trigger type
ammo type
noise level
fire rate
accuracy
mag size
max ammo
reload time
riven disposition
damage types
  this one how? slash, puncture, impact, heat, cold, toxin, shock, radiation, magnetic, viral, blast, 11 damage types total, have 11 columns? + 12ths for total damage per shot?
total damage
crit chance
crit multiplier
status chance
maaayybbe how to obtain? this would be hard to automate

at the end can output dps(rate of fire*dmg), dmg per mag? 


thenn the fun part, sorting data by damage types, fastest reload, most heat damage, best status, ect
output mastery level, sort by lowest up, ect

this basic template should work for all weapons, instance retrieved data as 1 row? in a largeish tibble
```{r setup}
library(rvest)
library(magrittr)
library(dplyr)
```

```{r}
#create iteratable list of urls to feed into scraper
u1 = 'https://warframe.fandom.com/wiki/'
```
```{r}
#grab weapon list from site scrape table?

```

``` {r braton test}
#read code from website
page1 <- read_html("https://warframe.fandom.com/wiki/Braton")
#table <- page1 %>%
  #html_nodes("table.pi-layout-default") %>%
  #html_table(header = F)
  #table <- table[5]

#without pipeline
table <- page1
links <- html_nodes(table, "pi-data-value pi-font")

x <- bind_rows(lapply(xml_attrs(links), function(x) data.frame(as.list(x), stringsAsFactors = FALSE)))

#test1 <- as.data.frame(table)

#table listed element from selector gadget isn't a table??? 

x
#arrrgggg how to get just a specific snippittttt out of the webpage without grabbing the 
#ENTIRE html block???
```















