---
title: "permitless carry  question"
author: "Jonathan Tan"
date: "9/1/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

dataset from kaggle -> https://www.kaggle.com/gunviolencearchive/gun-violence-database 
actually -> http://www.gunviolencearchive.org/reports 
lol dickey amendment 1997 so federal gun stats are illegal

req - 2 dataset interaction
- timeline of states that passed permitless carry laws (make distinction between open and concealed?)
- stats per year of those states? multiple sets to choose from in the dataset, accidental deaths, mass shootings, children / teens/ adults/ accidental injuries / deaths

hmm this dataset (kaggle) only hits the past 5 years? or is there more past data to find? 
take a look later at CDC numbers, other datasets? long term 20+ years would be ideal to match scope of permitless carry laws 

CDC firearm mortality by state - https://www.cdc.gov/nchs/pressroom/sosmap/firearm_mortality/firearm.htm 

get (make?) law timeline data for vis like wiki page - https://en.wikipedia.org/wiki/Constitutional_carry 
org 50ish columns = state, years on y with categorical gun law status
0 = no permitless carry
1 = permitless carry

usafacts.com has a compliation of past CDC data on firearm deaths
(dodgy name) (thirdparty bias check looks mostly factual though?)

```{r cars}
library(downloader)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(reshape2)
```
#arrange dataset in timeline / table by state / by category
```{r}
usaf <- read.csv('D:\\SMU\\Github Projects\\doodles\\gun_violence_dataset\\firearm_deaths_usafacts.csv')
#there we go 81 to 19, per state
#reorg usaf to select years and states only 
#reminder, r indexing is [y, x] 
fd1 <- usaf[27:77, 1:40]
#colnames(fd1) <- c("State", seq(1981, 2019)) #replace char with int
fd1[, 1] <- gsub('\\(People\\)', '', fd1[, 1]) %>% trimws()   #gsub escape is \\ 
fd1 <- fd1 %>% t() #transpose

colnames(fd1) <- fd1[1, ]
fd1 <- fd1[2:dim(fd1)[1], ] #bump state names to colnames, remove 1st col
rownames(fd1) <- seq(1981, 2019) #year to num
df1 <- data.frame(fd1)
df1 <- as.data.frame(sapply(df1, as.numeric)) #string to num
df1['years'] <- seq(1981, 2019)
```
```{r}
colnames(df1)
```
```{r}
df1[,1 ]
```
```{r}
#sanity vis
#year <- seq(1981, 2019)
state_num <- 1
g1 <- ggplot(df1, aes(years, df1[, state_num])) + geom_line() + ggtitle("CDC Firearm Deaths ", colnames(df1)[state_num])
g1
```

```{r}
#multiline graph requires melt
dfm <- melt(df1, id.vars = 'years' ,value.name = "deaths")
colnames(dfm) <- c('year', 'state', 'deaths')
g2 <- ggplot(dfm, aes(x = year, y = deaths, color = state)) + geom_line()
g2
```


#gun law timeline
from - https://en.wikipedia.org/wiki/Constitutional_carry 

```{r}
cc1 <- data.frame(matrix(data = 0, nrow = 39, ncol = 52))
colnames(cc1) <- colnames(df1)
cc1['year'] <- df1['years']
cc1 <- cc1[, c(52, 1:51)]
#now just need to put in 1's on the year that each state passes unrestricted carry? 
#maybe some degrees? some states have no permit open carry and no permit concealed carry, long guns vs hand guns? 
```



```{r}
c(52, 1:51)
```












